\bf Landslide: \\ Systematic Exploration for Kernel-Space Race Detection

### 2012-06-10 rv/paper.tex
final version of paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In kernel space, race condition debugging becomes even more difficult. Many aspects of the concurrency implementation itself are part of the system being tested.
While user-space testing frameworks operate above clear abstractions and interfaces, a systematic exporation tool for kernels cannot rely on such abstractions and interfaces.
Furthermore, execution in the scheduler and in similar components is common to all concurrent threads, even otherwise independent ones; this threatens our ability to achieve state-space reduction.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Among other systematic testing tools are CHESS \cite{chess}, with a heuristic for search ordering; MaceMC \cite{macemc} and MODIST \cite{modist}, which test networked and distributed applications; RacePRO \cite{racepro}, which uncovers inter-process races; and DeMeter \cite{demeter}, which focuses its search on only one component of a system at a time.

{\bf Kernel-level verification.}
% Kernel-level code has seen many dynamic testing techniques apart from systematic exploration.
DataCollider \cite{datacollider} uses memory access sampling to find data races in Windows 7. Compared to systematic exploration, data race detection is less expensive, but is limited in which types of races it can find.
% Carburizer \cite{carburizer} is a tool for enhancing resiliency of imperfect kernel drivers in the presence of faulty peripheral devices.
SimTester \cite{simtester} is another driver testing tool; like Landslide, it runs in Simics, and injects interrupts to control the system's concurrency. SimTester focuses on races involving device interrupts, whereas Landslide focuses on non-driver inter-thread races. Additionally, SimTester injects only one interrupt per test run, which is both less expensive and less powerful than systematic exploration.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scope}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Environment}
% 410, requirements, simulated

@@

To constrain the problem space some, we aimed to use a less complex kernel than production kernels such as Linux and Windows. We built Landslide to test Pebbles kernels, which students at CMU develop from scratch in a 6 week project. These kernels are necessarily less complex, yet still rife with race conditions that are debugged furiously every semester.
The Pebbles specification \cite{kspec} is a set of UNIX-like system calls which includes routines for thread lifecycle management and virtual memory management. Among these are \texttt{fork} and \texttt{exec}, \texttt{exit} (known as \texttt{vanish}) and \texttt{wait}, and \texttt{yield}.

{\bf Testing environment.} Landslide is implemented as a module for Simics, a full-system x86 simulator which students in 15-410 use as an execution environment for their kernels. Simics can load Landslide while booting a kernel, and calls into Landslide once per instruction and once per memory access of the simulated execution. Landslide will then perform systematic exploration on that kernel using Simics interface calls during the simulated execution.

Landslide assumes that timer interrupts are the only nondeterministic events that control a kernel's concurrent execution. While this prevents us from finding races related to external input such as disk or network I/O, Landslide is already able to find many types of complicated races by controlling timer-driven thread scheduling.
Landslide's model is also restricted to uniprocessor execution. Supporting multiprocessor environments, device driver testing, and production kernels is left to future work.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Challenges of Kernel Space}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIXME: maybe say in each paragraph how we address the issue?
% TODO: get rid of this section

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Design}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:design}
% prelude: define decision point and maybe voluntary reschedule

@@

Landslide automatically identifies a minimal set of decision points at voluntary reschedules between threads (e.g. \texttt{yield}). We provide an interface for configuring additional decision points (described in \S\ref{sec:config}).

\subsection{Exploration Mechanism}

@@

{\bf Thread scheduling.} Compared with user space testing frameworks, Landslide cannot rely on kernel-provided system calls to control the concurrent execution of the system under test. Instead, Landslide has an internal abstraction for scheduling arbitrary threads built upon injecting timer interrupts.
If thread $M$ is running and Landslide decides to preempt with thread $N$, it fires a timer interrupt and waits for the kernel to finish the triggered context switch. Then if the kernel is running a thread that is not $N$, Landslide repeats until thread $N$ begins running, at which point it lets the kernel continue execution.

@@

\subsection{Identifying Bugs}

@@

Landslide can optionally also check heuristically for infinite loops and livelock. This is done by comparing the length of the current interleaving against previously-explored ones: if the current interleaving has been running disproportionately longer (by a large arbitrary constant factor), it indicates the kernel is likely stuck.
We have never found this technique to erroneously report bugs (false positives), although it may fail to trigger when it should if too few previous interleavings have been tested to make a reliable comparison.

When Landslide identifies a bug, it halts execution and generates a {\em decision trace}. The trace reports what kind of bug was detected (for uses-after-free, it also prints stack traces for when the block was allocated and freed), and also reports each decision point in the current interleaving: which thread was running, its stack trace when it was switched away from, and the thread that we caused to preempt it. With this information, the user can better understand the concurrent execution that led to the race.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{User Interface}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:interface}

\subsection{Instrumenting Kernels with Landslide}
\label{sec:instrument}

@@

\subsection{Configuring Landslide}
\label{sec:config}

@@

This strategy may cause Landslide to identify decision points in unrelated parts of the kernel, such as when accessing mutexes in unrelated system calls. We provide interface options in \texttt{config.landslide} for the user to view currently identified decision points and to selectively limit them. First, the user writes ``\texttt{DECISION\_INFO\_ONLY}'' to make Landslide run only one interleaving and then print all decision points instead of continuing to explore. Then, the user writes ``\texttt{within\_function foo}'' to whitelist function \texttt{foo}, and ``\texttt{without\_function bar}'' to blacklist function \texttt{bar}. For example, if testing thread death and reaping, the user should write \texttt{within\_function exit}, and \texttt{within\_function wait}, and perhaps \texttt{without\_function destroy\_address\_space}.

{\bf Configuring Dynamic Partial Order Reduction.} Performing DPOR \cite{dpor} requires establishing a memory independence relation between transitions.
Unfortunately, since the kernel contains its own concurrency implementation, every thread switch will execute through a common path that modifies shared scheduler data structures.
Unchecked, this would cause every transition to conflict with each other and result in no reduction in execution time from DPOR.
Certain other shared memory conflicts also arise that are irrelevant to whichever system calls are being tested. For example, if testing for races in thread lifecycle routines, the user likely does not care about accesses to the frame allocator mutex.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% prob. don't actually say future work here; say it inline
% TODO

>>>>

% \newcommand\squish{\vspace{-8pt}}
\newcommand\squish{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\squish
\section{Introduction}
\squish
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In kernel space, race condition debugging becomes even more difficult. Many aspects of the concurrency implementation itself are part of the system being tested.
While user-space testing frameworks operate above clear abstractions and interfaces, a systematic exporation tool for kernels cannot rely on such easy means of understanding the system under test.
Furthermore, execution in the scheduler and in similar components is common to all concurrent threads, even otherwise independent ones; this threatens our ability to achieve state-space reduction.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\squish
\section{Related Work}
\squish
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Among other user-space systematic testing tools are CHESS \cite{chess}, with a heuristic for search ordering based on number of preemptions; MaceMC \cite{macemc} and MODIST \cite{modist}, which test networked and distributed applications; RacePRO \cite{racepro}, which uncovers inter-process races around system call invocations; and DeMeter \cite{demeter}, which focuses its search on only one component of a system at a time.

{\bf Kernel-level verification.}
% Kernel-level code has seen many dynamic testing techniques apart from systematic exploration.
DataCollider \cite{datacollider} uses memory access sampling to find data races in Windows 7. Compared to systematic exploration, data race detection is less expensive, but is limited in which types of races it can find.
% Carburizer \cite{carburizer} is a tool for enhancing resiliency of imperfect kernel drivers in the presence of faulty peripheral devices.
SimTester \cite{simtester} is a tool for testing kernel drivers; like Landslide, it runs in Simics, and injects interrupts to control the system's concurrency. SimTester focuses on races involving device interrupts, whereas Landslide focuses on non-driver inter-thread races. Additionally, SimTester injects only one interrupt per test run, which is less expensive than systematic exploration, but less powerful as well.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\squish
\section{Scope}
\squish
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Environment}
% 410, requirements, simulated

@@

To constrain the problem space some, we aimed to use a less complex kernel than production kernels such as Linux or Windows. We built Landslide to test Pebbles kernels, which students at CMU develop from scratch in a 6 week project. These kernels are necessarily less complex, yet still rife with race conditions that are debugged furiously every semester.
The Pebbles specification \cite{kspec} is a set of UNIX-like system calls which includes routines for thread lifecycle management and virtual memory management. Among these are \texttt{fork} and \texttt{exec}, \texttt{exit} (known as \texttt{vanish}) and \texttt{wait}, and \texttt{yield}.

{\bf Testing environment.} Landslide is implemented as a module for Simics, a full-system x86 simulator which students in 15-410 use as an execution environment for their kernels. Simics can load Landslide while booting a kernel, and calls into Landslide once per instruction and once per memory access of the simulated execution. Landslide will then perform systematic exploration on that kernel using Simics interface calls to control the kernel's execution.

Landslide assumes that timer interrupts are the only nondeterministic events that influence a kernel's concurrent execution. While this prevents us from finding races related to external input such as disk or network I/O, Landslide is already able to find many types of complicated races by controlling timer-driven thread scheduling.
Landslide's model is also restricted to uniprocessor execution. Supporting multiprocessor environments, device driver testing, and production kernels is left to future work.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Challenges of Kernel Space}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIXME: maybe say in each paragraph how we address the issue?

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\squish
\section{Design}
\squish
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:design}
% prelude: define decision point and maybe voluntary reschedule

@@

Landslide automatically identifies a minimal set of decision points at voluntary reschedules between threads (e.g. \texttt{yield}). We provide an interface for configuring additional decision points, described in \S\ref{sec:config}.

\squish
\subsection{Exploration Mechanism}
\squish

@@

{\bf Thread scheduling.} Compared with user-space testing frameworks, Landslide cannot rely on kernel-provided system calls to control the concurrent execution of the system under test. Instead, Landslide has an internal abstraction for scheduling arbitrary threads built upon injecting timer interrupts.
If thread $M$ is running and Landslide decides to preempt with thread $N$, it fires a timer interrupt and waits for the kernel to finish the triggered context switch. Then if the kernel is running a thread that is not $N$, Landslide repeats until thread $N$ begins running, at which point it lets the kernel continue execution.

@@

\squish
\subsection{Identifying Bugs}
\squish

@@

Landslide can optionally also check heuristically for infinite loops and livelock. This is done by comparing the length of the current interleaving against previously-explored ones: if the current interleaving has been running disproportionately longer (by a large arbitrary constant factor), it indicates the kernel is likely stuck.
We have never found this technique to erroneously report bugs (false positives), although it may fail to trigger when it ought to if too few previous interleavings have been tested to make a reliable comparison.

When Landslide identifies a bug, it halts execution and generates a {\em decision trace}. This trace reports what kind of bug was detected (for uses-after-free, it also prints stack traces for when the block was allocated and freed), and also reports each decision point in the current interleaving: which thread was running, a trace of its stack when it was switched away from, and the thread that we caused to preempt it. With this information, the user can better understand the concurrent execution that led to the race.

@@

%\begin{figure}
%	\centering
%	{\small
%	\begin{tabular}{l}
%	\texttt{void add\_to\_runqueue(tcb\_t *thread) \{} \\
%	\texttt{~~~~tell\_landslide\_on\_rq(thread->id);} \\
%	\texttt{~~~~list\_insert(\&runqueue, thread);} \\
%	\texttt{\}} \\
%	\end{tabular}
%	}
%	\caption{Example annotated code.}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\squish
\section{User Interface}
\squish
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:interface}

\subsection{Instrumenting Kernels with Landslide}
\squish
\label{sec:instrument}

@@

\squish
\subsection{Configuring Landslide}
\squish
\label{sec:config}

@@

This strategy may cause Landslide to identify decision points in unrelated parts of the kernel, such as when accessing mutexes in unrelated system calls. We provide interface options in \texttt{config.landslide} for the user to view currently identified decision points and to selectively limit them. First, the user writes ``\texttt{DECISION\_INFO\_ONLY}'' to make Landslide run only one interleaving and then print all decision points instead of continuing to explore. Then, the user writes ``\texttt{within\_function foo}'' to whitelist function \texttt{foo}, and ``\texttt{without\_function bar}'' to blacklist function \texttt{bar}. For example, if testing thread death and reaping, the user should write \texttt{within\_function exit}, and \texttt{within\_function wait}, and perhaps \texttt{without\_function destroy\_address\_space} (if they don't care about the virtual memory operations associated with \texttt{exit}).

{\bf Configuring Dynamic Partial Order Reduction.} Performing DPOR \cite{dpor} requires establishing a memory independence relation between transitions.
Unfortunately, since the kernel contains its own concurrency implementation, every thread switch will execute through a common path that modifies shared scheduler data structures.
Unchecked, this would cause every transition to conflict with each other and result in no reduction in execution time from DPOR.
Certain other shared memory conflicts also arise that are irrelevant to whichever system calls are being tested; for example, if testing for races in thread lifecycle routines, the user likely does not care about accesses to the frame allocator mutex.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\squish
\section{Evaluation}
\squish
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\squish
\section{Conclusion}
\squish
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We have built Landslide, a tool for systematic exploration on kernel-level code. Landslide runs as a Simics module during simulated execution of Pebbles kernels. It uses user-provided instrumentation to keep track of the kernel's state, such as the set of runnable threads, and it provides the user configuration options for making its search more effective. It identifies decision points in the kernel around which to interleave thread executions, and it controls how threads preempt each other by injecting timer interrupts. We found that Landslide identified previously unknown bugs in kernels we were already familiar with, and also that novice kernel programmers were able to find otherwise-overlooked bugs with Landslide without investing large amounts of time.
